Missing data: Current practice in football research and recommendations for

### 2020-10-04 paper/paper.Rmd
main text edits.

Our intention was to survey, at random, 10% of the total articles found (after 
title and abstract screening). We were primarily interested in
whether missing data was reported or acknowledged; and if not, whether
a dataset was shared. Where applicable, information relating to how missing
values were treated was also extracted (see Supplement 2). Two authors
independently extracted the data (X% agreement). Findings are reported
as the proportion and 95% confidence interval.

@@

Practice to avoid: *Unreported missing values.* While not always the case, even if
not reported, missing data can be obvious. A recent paper examined the
relationship between match performance indicators and outcome, in
Australian Football between the 2001 and 2016. Of the 91 team performance indicators included in the analysis,
at least one variable (i.e., 'meters gained') was not available over the entire 2001-2016 period [@Young2019]**Check this reference**. **We might need to add a sentence about 'question mark' in the figures?**
This was not noted anywhere in the paper, or in the supplementary
materials. It is unknown whether imputation was undertaken. Without
these details, it is difficult for other analysts and researchers to use
and/or extend the ideas presented in the paper, or reproduce the
analysis. It is also difficult for a reader to evaluate how the
missingness could bias the results. The absence of a
missing data statement when data are missing is not unique to the study
above [@Young2019]**Check this reference**. Authors should include a missing data statement,
irrespective of whether there is data missing, or not.

Practice to avoid: *Mean imputation.* The default of most statistical software for
handling missing data in modelling is listwise deletion. Listwise
deletion removes entire rows that contain missing observations from the
analysis. At best, listwise deletion reduces the power in analysis by
reducing the sample size. At worst, it introduces bias. For example, if
injured participants are removed from a study on injury prevention.
Imputing values removes the need to conduct listwise deletion. However,
it is critical how these values are imputed, and that the imputation
method is documented. One imputation method is imputing the mean value
(e.g., participant or group average) [@Young2019]**Check this reference**. While imputing the mean
preserves the study sample size (and point estimates), it also reduces
the variance, can alter the relationship between the variable with
missing observations and other variables, and can bias (underestimate)
standard errors [@Scheffer2002].
Smaller standard errors typically reduce *p*-values, which may lead to
incorrect inference. Mean imputation should be avoided. An imputation
strategy that incorporates information from other related variables in
the data -- such as linear regression, or K nearest neighbours -- should
be considered. This is discussed in the recommendations section.

Practice to avoid: *Not evaluating the effect, or choice, of imputation.* The 
method used to impute missing data has the potential to
affect the outcome of a study. It is important that authors
understand, and document, how imputation affected the goal of the
analysis (i.e., inference, prediction, or both). For example,
understanding the implications in using mean imputation, compared to
K-nearest neighbour imputation--in terms of the effect on parameter estimates
('significant' versus 'not significant'), and the (un)certainty of the
coefficients (i.e., less/more). When using imputation, authors need to
perform a sensitivity analysis (**ref circulation article-Nick**). This is
discussed below.

@@

Imputing data might feel wrong, as we are 'making up data'. The truth
is, we can (generally) never know what the missing values were. The goal
of imputing data is to make the best possible inference from the data.
We recommend data is imputed, with a few caveats. Consider removing
variables with a high proportion of missing data (e.g., variables with
the majority or more missing than not). We deliberately do not suggest a 'rule of
thumb' for the amount of missingness that should be imputed, and caution against
looking for a specific threshold. Rather, we recommend: values should generally be imputed, avoiding methods imputing the same value (such as the mean or median), and suggest using imputation methods such as: linear regression, k-nearest neighbours, or expectation
maximisation.

Missing values can occur in the both the predictors (the "independent variables") 
and response variable (the "dependent variables"). Care should be taken 
when imputing data, as typically only predictors should be imputed. 
For discussion on imputing the response value, see ... [@refs]. **(note: take a look again)**

@@

The current state of reporting missing data in the sports literature has room for improvement. In this section we discuss recommended practices for exploring and
reporting missing data. Our intention is not to provide a 'cookbook'
style approach to missing data, but rather, broad recommendations to help
researchers when writing the methods and results sections of a study,
and assist researchers when evaluating a study during the peer-review
process.

@@

> *25 of the 280 values in the dataset (~8.92%) were missing. These were due to faults in GPS tracking as participants ran through sections of track covered by forest, a known issue with some GPS tracking. These speed values were interpolated using a nearest neighbours approach, taking inputs of speed, and altitude. The track was imputed using the known track in the area for those sections.*

@@

Our survey of 2019 articles on football related topics
showed that current practice of reporting missing data is poor, 
with only about one in 20 studies reported missing data. 
This could suggest that this aspect of data anlaysis receives little attention. 
To address this issue and assist researchers,
we have provided recommendations for reporting and exploring missing data.
Researchers should consider these recommendations, and pay greater
attention to missing data and its influence on research results.

>>>>

Our intention was to survey, at random, 10% of the total articles found (after 
title and abstract screening). We were primarily interested in
whether missing data was reported or acknowledged; and if not, whether
a dataset was shared. Where applicable, information relating to how missing
values were treated was also extracted (see Supplement 2). Two authors
independently extracted the data (X% agreement). Results are reported
as the proportion and 95% confidence interval.

@@

Practice to avoid: *Unreported missing values.* While not always the case, even if
not reported, missing data can be obvious. For example, a recent study examined the
relationship between match performance indicators and outcome, in
Australian Football between the 2001 and 2016, where at least one of the
91 performance indicators ('meters gained') included in the analysis was not available over the entire 2001-2016 period [@Young2019]**Check this reference**. This was not noted anywhere in the paper, or in the supplementary
materials. It is unknown whether imputation was undertaken. Without
these details, it is difficult for other analysts and researchers to use
and/or extend the ideas presented in the paper, or reproduce the
analysis. It is also difficult for a reader to evaluate how the
missingness could bias the results. Authors should include a missing data statement,
irrespective of whether there is data missing, or not.

Practice to avoid: *Mean imputation.* The default of most statistical software for
handling missing data in modelling is listwise deletion. Listwise
deletion removes entire rows that contain missing observations from the
analysis. At best, listwise deletion reduces the power in analysis by
reducing the sample size. At worst, it introduces bias. For example, if
injured participants are removed from a study on injury prevention.
Imputing values removes the need to conduct listwise deletion. However,
it is critical how these values are imputed, and that the imputation
method is documented. Authors sometimes choose to replace missing 
observations with mean values (e.g., participant or group average) [@refs](https://doi.org/10.1016/j.jsams.2018.05.019, https://doi.org/10.1016/j.jsams.2008.01.009). 
While imputing the mean preserves the study sample size (and point estimates), it also reduces
the variance, can alter the relationship between the variable with
missing observations and other variables, and can bias (underestimate)
standard errors [@Scheffer2002]. Smaller standard errors typically reduce *p*-values, which may lead to
incorrect inference. Mean imputation should be avoided. An imputation
strategy that incorporates information from other related variables in
the data -- such as linear regression, or K nearest neighbours -- should
be considered. This is discussed in the recommendations section.

Practice to avoid: *Not evaluating the effect, or choice, of imputation.* The 
method used to impute missing data has the potential to
affect the outcome of a study. It is important that authors
understand, and document, how imputation affected the goal of the
analysis (i.e., inference, prediction, or both). For example,
understanding the implications in using mean imputation, compared to
K-nearest neighbour imputation--in terms of the effect on parameter estimates
('significant' versus 'not significant'), and the (un)certainty of the
coefficients (i.e., less/more). When using imputation, authors need to
perform a sensitivity analysis (**ref circulation article-Nick, possible reference from dave - https://doi.org/10.1191%2F1740774504cn032oa**). This is
discussed below.

@@

Imputing data might feel wrong, as we are 'making up data'. The truth
is, we can (generally) never know what the missing values were. The goal
of imputing data is to make the best possible inference from the data.
We recommend data is imputed, with a few caveats. Consider removing
variables with a high proportion of missing data (e.g., variables with
the majority or more missing than not). We deliberately do not suggest a 'rule of
thumb' for the amount of missingness that should be imputed, and caution against
looking for a specific threshold. Rather, we recommend: values should generally be imputed, avoiding methods imputing the same value (e.g., the mean or median), and suggest using imputation methods such as: linear regression, k-nearest neighbours, or expectation
maximisation.

Missing values can occur in the both the predictors (the "independent variables") 
and response variable (the "dependent variables"). Care should be taken 
when imputing data, as typically only predictors should be imputed. 
For discussion on imputing the response value, see ... [@refs]. **(note: take a look again - this is the discusson of missing y values section - and possible reference relevant to RCT's is this: https://doi.org/10.1191%2F1740774504cn032oa)**

@@

The current state of reporting missing data in football research has room for improvement. 
In this section we discuss recommended practices for exploring and
reporting missing data. Our intention is not to provide a 'cookbook'
style approach to missing data, but rather, broad recommendations to help
authors when writing the methods and results sections of a study,
and assist researchers when evaluating a study during the peer-review process.

@@

> *Twenty-five of the 280 values in the dataset (~8.92%) were missing. These were due to faults in GPS tracking as participants ran through sections of track covered by forest, a known issue with some GPS tracking. These speed values were interpolated using a nearest neighbours approach, taking inputs of speed, and altitude. The track was imputed using the known track in the area for those sections.*

@@

Our survey of 2019 articles on football related topics
showed that current practice of reporting missing data is poor, 
with only about one in 20 studies reporting missing data.
We speculate that this result is representative of practice in the 
wider sports science literature. The extremely low reporting of missing 
data could suggest that this aspect of data anlaysis receives little attention. 
To address this issue and assist researchers,
we have provided recommendations for reporting and exploring missing data.
Researchers should consider these recommendations, and pay greater
attention to missing data and its influence on research results.
